---
title: "Fairness analysis"
author: "magica"
date: "10/09/2018"
output: html_document
---

# Function definitions, setup

```{r}
library(stringr)
library(ggplot2)
library(dplyr)
library(magrittr)
library(corrplot)
library(robust)
library(ggpubr)
library(fit.models)
library(matrixStats)
```

We'll be showing many different charts with a large number of different attributes, so a good
categorical color scale is helpful. We're using [d3](https://d3js.org) `d3.schemeCategory10` and `d3.schemeCategory20`.

```{r}
# Use these for more than 10 algorithms
colors_20 = c(
  "#1f77b4", "#aec7e8", "#ff7f0e", "#ffbb78", "#2ca02c", "#98df8a", "#d62728", "#ff9896",
  "#9467bd", "#c5b0d5", "#8c564b", "#c49c94", "#e377c2", "#f7b6d2", "#7f7f7f", "#c7c7c7",
  "#bcbd22", "#dbdb8d", "#17becf", "#9edae5")

# Use these for 10 or fewer algorithms
colors_10 = c("#1f77b4", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd", "#8c564b", "#e377c2", "#7f7f7f", "#bcbd22", "#17becf")

symbol = c(21:25, 1, 0, 5, 2, 6,
           7:14, 3, 4)

catscale10  = scale_colour_manual(values=colors_10)
catscale10_2 = scale_fill_manual(values=colors_10)

catscale20  = scale_colour_manual(values=colors_20)
catscale20_2 = scale_fill_manual(values=colors_20)

algos_to_plot = c(
  "UGP_dem_par_in_True_tar_0.1", "UGP_dem_par_in_True_tar_0.2", "UGP_dem_par_in_True_tar_0.3", "UGP_dem_par_in_True_tar_0.4", "UGP_dem_par_in_True_tar_0.5",
  "UGP_dem_par_in_False_tar_0.1", "UGP_dem_par_in_False_tar_0.2", "UGP_dem_par_in_False_tar_0.3", "UGP_dem_par_in_False_tar_0.4", "UGP_dem_par_in_False_tar_0.5"
)

make_scatter_figure = function(name, var1="race-TPRDiff", var2="accuracy", algos=algos_to_plot, display="off", statistic = "off") {
  x_var = as.name(var1)
  y_var = as.name(var2)
  
  df = read.csv(str_c(name, "_numerical-binsensitive.csv"), check.names=FALSE) %>%
    filter(algorithm %in% algos) %>%
    # mutate(algorithm=recode(algorithm, ZafarFairness="ZafarFairness")) # rename to Zafar for clarity
    mutate(algorithm=recode(algorithm, "UGP_dem_par_in_True_tar_0.1"="FairGPparity*, target=0.1"))%>%
    mutate(algorithm=recode(algorithm, "UGP_dem_par_in_True_tar_0.2"="FairGPparity*, target=0.2"))%>%
    mutate(algorithm=recode(algorithm, "UGP_dem_par_in_True_tar_0.3"="FairGPparity*, target=0.3"))%>%
    mutate(algorithm=recode(algorithm, "UGP_dem_par_in_True_tar_0.4"="FairGPparity*, target=0.4"))%>%
    mutate(algorithm=recode(algorithm, "UGP_dem_par_in_True_tar_0.5"="FairGPparity*, target=0.5"))%>%
    mutate(algorithm=recode(algorithm, "UGP_dem_par_in_False_tar_0.1"="FairGPparity, target=0.1"))%>%
    mutate(algorithm=recode(algorithm, "UGP_dem_par_in_False_tar_0.2"="FairGPparity, target=0.2"))%>%
    mutate(algorithm=recode(algorithm, "UGP_dem_par_in_False_tar_0.3"="FairGPparity, target=0.3"))%>%
    mutate(algorithm=recode(algorithm, "UGP_dem_par_in_False_tar_0.4"="FairGPparity, target=0.4"))%>%
    mutate(algorithm=recode(algorithm, "UGP_dem_par_in_False_tar_0.5"="FairGPparity, target=0.5"))
  
  switch (statistic,
          median = {
            new_df = matrix(nrow=N_algos,ncol=2, dimnames =list(1:N_algos, c(var1, var2)))
            for(j in 1:N_algos){
              temp_data = df[which(df$algorithm==algos[j]),c(var1, var2)]
              new_df[j, ] =  colMedians(data.matrix(temp_data))
            }
            dataFrame <- data.frame(new_df)
            names(dataFrame) <- c(var1,var2)
            dataFrame$algorithm <- algos
          },
          mean = {
            new_df = matrix(nrow=N_algos,ncol=2, dimnames =list(1:N_algos, c(var1, var2)))
            for(j in 1:N_algos){
              temp_data = df[which(df$algorithm==algos[j]),c(var1, var2)]
              new_df[j, ] =  colMeans(data.matrix(temp_data))
            }
            dataFrame <- data.frame(new_df)
            names(dataFrame) <- c(var1,var2)
            dataFrame$algorithm <- algos
          },
          off = {
            dataFrame = df
          }
          
  )
  # Scatterplot
  color_used = catscale10
  color_used_2 = catscale10_2
  target_view = as.name("algorithm")
  
  if(grepl("sex", var1)){
    var1 = gsub('sex', 'gender', var1)
  }
  if(grepl("sex", var2)){
    var2 = gsub('sex', 'gender', var2)
  }
  if(grepl("sex", name)){
    name = gsub('sex', 'gender', name)
    fig_title = "Adult (Gender)"
  }else{
    fig_title = "Adult (Race)"
  }
  
  
  switch(display,
         off = ggplot(dataFrame, aes_q(x=x_var, y=y_var, colour=target_view, bg=target_view)) + color_used + color_used_2 +
           geom_point(size=3, aes_q(shape=target_view)) + scale_shape_manual(values=symbol) + theme_bw() + 
           labs(y=var2, x=var1, title=fig_title),
         title = ggplot(dataFrame, aes_q(x=x_var, y=y_var, colour=target_view, bg=target_view)) + color_used + color_used_2 +
           geom_point(size=3, aes_q(shape=target_view)) + scale_shape_manual(values=symbol) + theme_bw() + 
           labs(y=var2, x=var1, title=fig_title),
         captain = ggplot(dataFrame, aes_q(x=x_var, y=y_var, colour=target_view, bg=target_view)) + color_used + color_used_2 +
           geom_point(size=3, aes_q(shape=target_view)) + scale_shape_manual(values=symbol) + theme_bw() + 
           labs(y=var2, x=var1, caption = str_c(name, "_numerical-binsensitive.csv")),
         all = ggplot(dataFrame, aes_q(x=x_var, y=y_var, colour=target_view, bg=target_view)) + color_used + color_used_2 +
           geom_point(size=3, aes_q(shape=target_view)) + scale_shape_manual(values=symbol) + theme_bw() + 
           labs(y=var2, x=var1, title=fig_title)
  )
}

```

## Users setting

```{r Users setting}
file_name = list("adult_race", "adult_sex")
file_attribute = list("race", "sex")
N_file = length(file_name)

```

## Algorithm accuracy vs fairness (DIbinary)

```{r}
variable_1 = "DIbinary"
add_for_1 = 0
target = "accuracy"
add_for_target = 0

plot_for_in  = c(
  "UGP_dem_par_in_True_tar_0.1", "UGP_dem_par_in_True_tar_0.2", "UGP_dem_par_in_True_tar_0.3", "UGP_dem_par_in_True_tar_0.4", "UGP_dem_par_in_True_tar_0.5"
)
plot_for = c(
  "UGP_dem_par_in_False_tar_0.1", "UGP_dem_par_in_False_tar_0.2", "UGP_dem_par_in_False_tar_0.3", "UGP_dem_par_in_False_tar_0.4", "UGP_dem_par_in_False_tar_0.5"
)

for (i in 1:N_file) {
  if(add_for_target){
    target_name = paste(file_attribute[i], "-", target, sep="")
  } else {
    target_name = target
  }
  if(add_for_1){
    var1_name = paste(file_attribute[i], "-", variable_1, sep="")
  } else {
    var1_name = variable_1
  }
  q = make_scatter_figure(file_name[i], var1=var1_name, var2=target_name)
  print(q)
  export_name = paste(file_name[i], "_tar.eps", sep="")
  ggsave(export_name, q)
}

```

## Algorithm 0-PR vs 1-PR

```{r}

variable_1 = "0-PR"
variable_2 = "1-PR"

for (i in 1:N_file) {
  var1_name = paste(variable_1, "-", file_attribute[i], sep="")
  var2_name = paste(variable_2, "-", file_attribute[i], sep="")
  
  # q_1 = make_scatter_figure(file_name[i], var1=var1_name, var2 = var2_name, algos=plot_for_in, display = "all")
  # print(q_1)
  # export_name = paste(file_name[i], "_PR_in.eps", sep="")
  # ggsave(export_name, q_1)
  
  p = make_scatter_figure(file_name[i], var1=var1_name, var2=var2_name)
  print(p)
  export_name = paste(file_name[i], "_PR.eps", sep="")
  ggsave(export_name, p)
}
```

## box plot

```{r box plot}
# plot box
create_sensitivity_summary = function(df, x_var, y_var) {
  quo_x_var = enquo(x_var)
  quo_y_var = enquo(y_var)
  
  
  sensitivity_summary =
    df %>%
    group_by(algorithm) %>%
    summarise(x_sd=sd(!!quo_x_var), x_mean=mean(!!quo_x_var),
              y_sd=sd(!!quo_y_var), y_mean=mean(!!quo_y_var))
  sensitivity_summary
}

basic_sensitivity_plot = function(sensitivity_summary, x_var, y_var) {
  if(grepl("sex", quo_name(enquo(x_var)))){
    label_x = gsub('sex', 'gender', quo_name(enquo(x_var)))
  } else {
    label_x = quo_name(enquo(x_var))
  }
  
  if(grepl("sex", quo_name(enquo(y_var)))){
    label_y = gsub('sex', 'gender', quo_name(enquo(y_var)))
  } else {
    label_y = quo_name(enquo(y_var))
  }
  
  ggplot(sensitivity_summary, aes(x=x_mean, y=y_mean, colour=algorithm)) + catscale20 + catscale20_2 +
    xlab(quo_name(enquo(x_var))) + ylab(quo_name(enquo(y_var)))
}

plot_lines_sensitity_summary = function(basic_plot) {
  basic_plot +
    geom_segment(aes(x=x_mean - x_sd, xend = x_mean + x_sd,
                     y=y_mean, yend = y_mean,
                     colour=algorithm)) +
    geom_segment(aes(x=x_mean, xend = x_mean,
                     y=y_mean - y_sd, yend = y_mean + y_sd,
                     colour=algorithm))
}

plot_ellipses_sensitity_summary = function(basic_plot) {
  basic_plot + stat_ellipse(level=0.5, geom="polygon", aes(fill=algorithm), alpha=0.2) + stat_ellipse(level=0.5, fill=NA)
}

plot_rects_sensitivity_summary = function(basic_plot) {
  aesthetics1 = aes(xmin = x_mean - 0.5 * x_sd,   
                    xmax = x_mean + 0.5 * x_sd,
                    ymin = y_mean - 0.5 * y_sd, 
                    ymax = y_mean + 0.5 * y_sd,
                    fill=algorithm)
  aesthetics2 = aes(xmin = x_mean - 0.5 * x_sd,   
                    xmax = x_mean + 0.5 * x_sd,
                    ymin = y_mean - 0.5 * y_sd, 
                    ymax = y_mean + 0.5 * y_sd,
                    colour=algorithm)
  basic_plot + theme_bw() + geom_point(aes(x_mean, y_mean), size = 2) + 
    geom_rect(aesthetics1, alpha=0.5, size=1.5) + 
    geom_rect(aesthetics2, fill=NA)
  
}

do_it_all = function(name, var1, var2, check.names=FALSE) {
  df = read.csv(str_c(name, ".csv"), check.names=check.names)
  
  x_var = as.name(var1)
  y_var = as.name(var2)
  
  sensitivity_summary = create_sensitivity_summary(df, !!x_var, !!y_var)
  
  basic_plot = basic_sensitivity_plot(sensitivity_summary, !!x_var, !!y_var) + ggtitle(name)
  
  plot_rects_sensitivity_summary(basic_plot)
}

nb_measure_comparison = function(name, measure) {
  num <- read.csv(str_c(name, "_numerical.csv"), check.names=FALSE)
  nbs <- read.csv(str_c(name, "_numerical-binsensitive.csv"), check.names=FALSE)
  
  var1 <- as.name(str_c(measure, "-numerical"))
  var2 <- as.name(str_c(measure, "-binsensitive"))
  
  df <- inner_join(num, nbs, by=c("algorithm", "run-id"), suffix = c("-numerical", "-binsensitive"))
  df$name = name
  df
  # return (df %>% select(!!var1, !!var2, "run-id", "algorithm", "name"))
}

plot_sensitivity = function(df, var1, var2) {
  x_var = as.name(var1)
  y_var = as.name(var2)
  
  sensitivity_summary = create_sensitivity_summary(df, !!x_var, !!y_var)
  basic_plot = basic_sensitivity_plot(sensitivity_summary, !!x_var, !!y_var)
  plot_rects_sensitivity_summary(basic_plot)
}

make_sensitivity_figure = function(name, var1="DIbinary", var2="accuracy", algos_plot=algos_to_plot) {
  df = read.csv(str_c(name, "_numerical-binsensitive.csv"), check.names=FALSE) %>%
    filter(algorithm %in% algos_plot) %>%
    # mutate(algorithm=recode(algorithm, ZafarFairness="Zafar")) # rename to Zafar for clarity
    mutate(algorithm=recode(algorithm, "UGP_dem_par_in_True_tar_0.1"="FairGPparity*, target=0.1"))%>%
        mutate(algorithm=recode(algorithm, "UGP_dem_par_in_False_tar_0.1"="FairGPparity, target=0.1"))%>%
    mutate(algorithm=recode(algorithm, "UGP_dem_par_in_True_tar_0.2"="FairGPparity*, target=0.2"))%>%
        mutate(algorithm=recode(algorithm, "UGP_dem_par_in_False_tar_0.2"="FairGPparity, target=0.2"))%>%
    mutate(algorithm=recode(algorithm, "UGP_dem_par_in_True_tar_0.3"="FairGPparity*, target=0.3"))%>%
        mutate(algorithm=recode(algorithm, "UGP_dem_par_in_False_tar_0.3"="FairGPparity, target=0.3"))%>%
    mutate(algorithm=recode(algorithm, "UGP_dem_par_in_True_tar_0.4"="FairGPparity*, target=0.4"))%>%
        mutate(algorithm=recode(algorithm, "UGP_dem_par_in_False_tar_0.4"="FairGPparity, target=0.4"))%>%
    mutate(algorithm=recode(algorithm, "UGP_dem_par_in_True_tar_0.5"="FairGPparity*, target=0.5"))%>%
        mutate(algorithm=recode(algorithm, "UGP_dem_par_in_False_tar_0.5"="FairGPparity, target=0.5"))
  
  df$algorithm = factor(df$algorithm, levels= order) 
  if(grepl("sex", name)){
    # name = gsub('sex', 'gender', name)
    namefile = "Adult (Gender)"
    plot_sensitivity(df, var1, var2)  + ggtitle(namefile) # + xlim(0.75, 0.95)
  }else{
    namefile = "Adult (Race)"
    plot_sensitivity(df, var1, var2)  + ggtitle(namefile) # + xlim(1.05, 1.25)
  }
}
```

## Algorithm sensitivity variable_1 vs variable_2

```{r}
order = c(
  "FairGPparity*, target=0.1", "FairGPparity, target=0.1",
  "FairGPparity*, target=0.2", "FairGPparity, target=0.2",
  "FairGPparity*, target=0.3", "FairGPparity, target=0.3",
  "FairGPparity*, target=0.4", "FairGPparity, target=0.4",
  "FairGPparity*, target=0.5", "FairGPparity, target=0.5"
)

variable_1 = "DIbinary"
add_for_1 = 0

variable_2 = "accuracy"
add_for_2 = 0

for (i in 1:N_file) {
  # temp_name = paste(file_name[i], "_numerical-binsensitive.csv", sep="")
  if(add_for_1){
    var1_name = paste(file_attribute[i], "-", variable_1, sep="")
  } else {
    var1_name = variable_1
  }
  if(add_for_2){
    var2_name = paste(file_attribute[i], "-", variable_2, sep="")
  } else {
    var2_name = variable_2
  }
  p = make_sensitivity_figure(file_name[i], var1 = var1_name, var2 = var2_name)
  print(p)
  export_name = paste(file_name[i], "_tar.eps", sep="")
  #ggsave(export_name, p)
}

```

## Algorithm Algorithm 0-PR vs 1-PR

```{r}
variable_1 = "0-PR"
variable_2 = "1-PR"

for (i in 1:N_file) {
  # temp_name = paste(file_name[i], "_numerical-binsensitive.csv", sep="")
  var1_name = paste(variable_1, "-", file_attribute[i], sep="")
  var2_name = paste(variable_2, "-", file_attribute[i], sep="")
  
  p = make_sensitivity_figure(file_name[i], var1 = var1_name, var2 = var2_name)
  print(p)
  export_name = paste(file_name[i], "_PR.eps", sep="")
  #ggsave(export_name, p)
}
```